### 2016-07-04 case-studies.tex
edits to bib and wording changes in Sec. 3 and 6

In this section, we present two case studies that demonstrate the
applicability of our reasoning approach in varied settings, and its
consequent practical utility. Both our case studies target
Ruby-on-Rails applications. To analyze these, and other similar
open-source applications, we developed a symbolic execution engine
called {\sc MagLev} that compiles a Rails program to an abstract
program in an extended version of \txnimp\footnote{The details of {\sc
MagLev} can be found in the supplementary. Its source code, along with
that of our applications and experiments, is available at the
following anonymized url: \url{...}}. The focus of our formal
verification efforts are such \txnimp programs.

\paragraph{A video view counter.} Our first application is a Rails
implementation of a counter to count the number of views a video has
garnered on a video-sharing website like YouTube. The application
stores the video count in a MySQL database, and supports a read
operation (a database read transaction), and an increment operation (a
read-increment-write sequence wrapped in a transaction).  The
application is expected to preserve the \emph{monotonicity invariant},
namely that the view count should never appear to be decreasing.
However, the default \iso{Repeatable Read} isolation level of MySQL
leads to anomalous executions - with 32 concurrent writers performing
single increments, and one reader issuing a constant stream of reads,
we observed 12 violations of monotonicity in 10 rounds. We
subsequently instantiated our reasoning framework for an {\sc sc}
store with ANSI SQL isolation levels, and spent 5 man hours
determining appropriate isolation levels needed to enforce the
invariant. In particular, we were able to formally prove that to
preserve the invariant, writers need to execute using
\iso{Serializable} isolation, while readers can execute using
\iso{Read Uncommitted}. Repeating the experiments with this
configuration led to no violations, as expected. While executing all
transactions at {\sc ser} also prevents violations, it led to an
additional 113\% increase in the latency of writes averaged over 5
rounds with 32 concurrent writers.  We repeated the experiments with
the same configuration on Postgres, and found no violations. However,
Postgres' isolation levels~\cite{postgresiso} are stronger than ANSI
SQL's in non-trivial ways. We therefore formalized Postgres'
guarantees and repeated relevant parts of the proof to discover that
\iso{Repeatable Read} is in fact sufficient for writers - an
observation we confirmed through experiments. As usual, using {\sc
ser} for all transactions also suffices, but on Postgres results in an
additional 47\% increase in write latency.

@@

\paragraph{Microblog.} Our second set of experiments focused on a
Twitter-like microblogging application taken from~\cite{railsbook}.
An important invariant that must be preserved is the referential
integrity between microposts and authors. The application enforces
this invariant entirely in user space by relying on database-backed
transactions (this behaviour is typical of Rails
applications~\cite{bailisferal}).  For instance, deletion of a user is
carried out in a transaction that also includes the deletion of
microposts authored by the user.  However, it is possible for
concurrent observers to nonetheless witness invariant violations. To
measure the likelihood of such anomalies, we first populated a
Postgres database with 1000 user accounts, each user with 50
microposts and 20 followers, chosen uniformly at random. We then
constructed experiments with 64 concurrent clients performing
deletions and timeline reads of randomly chosen users in a 1:7 ratio,
and witnessed 250 violations of referential integrity (the isolation
level was left at the default {\sc rc} level for experiments). To
determine appropriate isolation levels, we then instantiated our
framework for an {\sc sc} store with various ANSI SQL isolation levels
for user deletions and timeline reads, and obtained a proof that {\sc
rc} is sufficient for deletions, whereas reads require {\sc rr}. The
proof is immediately applicable for MySQL, which implements ANSI SQL
isolation levels. Repeating our experiments for MySQL with this
configuration resulted in no violations (as expected), while yielding
a 38\% reduction in latency than an {\sc ser} configuration.  Since
Postgres's isolation levels are stronger than the corresponding ANSI
levels, \rulelabel{RG-Conseq} rule (Fig.~\ref{fig:rg-rules}) lets us
immediately deduce that the same choice of isolation levels also work
on Postgres.

Clearly, there exists a significant diversity among the weak
consistency and isolation variants implemented on various real-world
data stores. More such variants have been developed in recent years to
meet the need for high availability (\emph{e.g.}, \emph{Parallel
  Snapshot Isolation}~\cite{psi} (2011) or \emph{Non-Monotonic
  Snapshot Isolation}~\cite{nmsi} (2013)).  As more commercial
weakly-consistent stores adopt various kinds of transactional
semantics, like Cassandra's recently introduced \emph{lightweight
  transactions}, it is reasonable to expect new isolation
definitions to continue to be proposed.  While it may be possible to
carefully engineer a reasoning framework for each combination of
consistency and isolation, such a strategy would lead to multiple
semantic definitions and proof systems with no obvious way to compare
and relate them. As illustrated here, having a parameterized semantics
and a proof framework built on top of it allows us to support these
variants as distinct, yet comparable, instantiations.

>>>>

In this section, we present two case studies that demonstrate the
applicability of our framework in varied settings, and its consequent
practical utility. Both our case studies target Ruby-on-Rails
applications. To analyze these, and other similar open-source
applications, we have developed a symbolic execution engine that
compiles a Rails program to an abstract program in an extended version
of \txnimp, along with some proof automation support.\footnote{Details
  ( can be found in the supplementary material.  Source code for our
  implementation, along with applications and experiments, are
  available at the following anonymized url: \url{...}} The focus of
our formal verification efforts are such translated \txnimp programs.

\paragraph{A video view counter.} Our first application is a Rails
implementation of a counter to count the number of views a video has
garnered on a video-sharing website like YouTube. The application
stores the video count in a MySQL database, and supports a read
operation (a database read transaction), and an increment operation (a
read-increment-write sequence wrapped in a transaction).  The
application is expected to preserve a \emph{monotonicity invariant},
namely that the view count should never appear to be decreasing.
However, the default \iso{Repeatable Read} isolation level of MySQL
leads to anomalous executions - with 32 concurrent writers performing
single increments, and one reader issuing a constant stream of reads,
we observed 12 violations of monotonicity in 10 rounds. We
subsequently instantiated our reasoning framework for an {\sc sc}
store with ANSI SQL isolation levels, and spent 5 man hours
determining appropriate isolation levels needed to enforce the
invariant. In particular, we were able to formally prove that to
preserve the invariant, writers need to execute using
\iso{Serializable} isolation, while readers can execute using
\iso{Read Uncommitted}. Repeating the experiments with this
configuration led to no violations, as expected. While executing all
transactions at {\sc ser} also prevents violations, it led to an
additional 113\% increase in the latency of writes averaged over 5
rounds with 32 concurrent writers.  We repeated the experiments with
the same configuration on Postgres, and found no violations. However,
Postgres' isolation levels~\cite{postgresiso} are stronger than ANSI
SQL's in non-trivial ways. We therefore formalized Postgres'
guarantees and repeated relevant parts of the proof to discover that
\iso{Repeatable Read} is in fact sufficient for writers - an
observation further validated through experiments. As usual, using
{\sc ser} for all transactions also suffices, but on Postgres results
in an additional 47\% increase in write latency.

@@

\paragraph{Microblog.} Our second set of experiments focused on a
Twitter-like microblogging application taken from~\cite{railsbook}.
An important invariant that must be preserved is referential integrity
between microposts (tweets) and authors. The application enforces this
invariant entirely in user space by relying on database-backed
transactions (this behaviour is typical of Rails
applications~\cite{bailisferal}).  For instance, the deletion of a
user is carried out in a transaction that also includes the deletion
of microposts authored by the user.  However, it is possible for
concurrent observers to nonetheless witness invariant violations. To
quantify the likelihood of such anomalies, we first populated a
Postgres database with 1000 user accounts, in which each user is
associated with 50 microposts and 20 followers, chosen uniformly at
random. We then constructed experiments with 64 concurrent clients
performing deletions and timeline reads of randomly chosen users in a
1:7 ratio, and witnessed 250 violations of referential integrity (the
isolation level was left at the default {\sc rc} level for these
experiments). To determine appropriate (correct) isolation levels, we
then instantiated our framework for an {\sc sc} store with various
ANSI SQL isolation levels for user deletions and timeline reads, and
obtained a proof that {\sc rc} is sufficient for deletions, whereas
reads require {\sc rr}. The proof is immediately applicable for MySQL,
which implements ANSI SQL isolation levels. Repeating our experiments
for MySQL with this configuration resulted in no violations (as
expected), while yielding a 38\% reduction in latency than an {\sc
  ser} configuration.  Since Postgres' isolation levels are stronger
than the corresponding ANSI levels, \rulelabel{RG-Conseq} rule
(Fig.~\ref{fig:rg-rules}) lets us immediately deduce that the same
choice of isolation levels also work on Postgres.

Clearly, there exists significant diversity among the weak consistency
and isolation variants implemented on various real-world data
stores. More such variants have been developed in recent years to meet
the need for high availability (\emph{e.g.}, \emph{Parallel Snapshot
  Isolation}~\cite{psi} (2011) or \emph{Non-Monotonic Snapshot
  Isolation}~\cite{nmsi} (2013)).  As more commercial
weakly-consistent stores adopt various kinds of transactional
semantics, like Cassandra's recently introduced \emph{lightweight
  transactions}, it is reasonable to expect new isolation definitions
to continue to be proposed.  While it may be possible to carefully
engineer a reasoning framework for each combination of consistency and
isolation, such a strategy would lead to multiple semantic definitions
and proof systems with no obvious way to compare and relate them. As
illustrated here, having a parameterized semantics and a proof
framework built on top of it allows us to support these variants as
distinct, yet comparable, instantiations.

### 2016-07-04 motivation.tex
edits to bib and wording changes in Sec. 3 and 6

Fig.~\ref{fig:motiv-eg-1-hb} is a visualization of $\hbZ$ from \C{Wd2}
to \C{Wd1} allowed by $\psi_{RC}$ and $\psi_{SI}$. All arrows are
legal under $\psi_{RC}$ because in no case does an operation $\eta_1$
from \C{Wd1} happen-before $\eta_2$ of \C{Wd2} without the commit of
\C{Wd1} also happening before $\eta_2$. $\psi_{SI}$ however disallows
all dashed arrows, because dashed arrows establish $\hbZ$ between
actions in \C{Wd2} and only a subset of the operations in \C{Wd1}.
Since the result of an operation (\eg a read) depends on what happens
before that operation, $\hbZ$ relation ultimately affects the result
of the program. The solid arrow and the first (from left) dashed arrow
in Fig.~\ref{fig:motiv-eg-1-hb} denote $\hbZ$ relationships that do
not affect the value of \C{C} in a way that causes the program to
violate its post-condition. The last two dashed arrows denote $\hbZ$
relationships that lead to the violation of the post-condition. The
task of the reasoning framework is to determine if all $\hbZ$
relationships allowed by an isolation level lead to the satisfaction
of the post-condition.

>>>>

Fig.~\ref{fig:motiv-eg-1-hb} is a visualization of $\hbZ$ from \C{Wd2}
to \C{Wd1} allowed by $\psi_{RC}$ and $\psi_{SI}$. All arrows are
legal under $\psi_{RC}$ because in no case does an operation $\eta_1$
from \C{Wd1} happen-before $\eta_2$ of \C{Wd2} without the commit of
\C{Wd1} also happening before $\eta_2$. $\psi_{SI}$ however disallows
all dashed arrows, because dashed arrows establish $\hbZ$ edges
between actions in \C{Wd2} and only a subset of the operations in
\C{Wd1}.  Since the result of an operation (\eg a read) depends on
what happens before that operation, the structure of the $\hbZ$
relation ultimately dictates the final result of the program. The
solid arrow and the first (from left) dashed arrow in
Fig.~\ref{fig:motiv-eg-1-hb} denote $\hbZ$ relationships that do not
affect the value of \C{C} in a way that causes the program to violate
its post-condition. The last two dashed arrows denote $\hbZ$
relationships that lead to the violation of the post-condition. The
task of the reasoning framework is to determine if all $\hbZ$
relationships allowed by an isolation level lead to the satisfaction
of the post-condition.
